---
title: "存活分析_HW1"
author: "B082040005 高念慈"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=F, results='hide', message=FALSE, warning=FALSE}

# install.packages('survival')
# install.packages('KMsurv')
# install.packages("fastDummies")
library(survival)                # survival function
library(KMsurv)                  # 資料集

# library(fastDummies)             # dummy variable

```

### data(bfeed):餵母乳

### Description
The bfeed data frame has 927 rows and 10 columns.<br>

Format<br>
This data frame contains the following columns:<br>

+ Duration: Duration of breast feeding, weeks<br>
+ delta: Indicator of completed breast feeding (1=yes, 0=no)
+ race: Race of mother (1=white, 2=black, 3=other)
+ poverty(貧窮): Mother in poverty (1=yes, 0=no)
+ yschool: Education level of mother (years of school)

We would like to investigate(調查) the relation <br>
between duration of breast feeding weeks and several covariates, <br>
including race of mother (race), whether the mother is in poverty (poverty), <br>
and mother’s education level (yschool).

```{r}
data(bfeed)
head(bfeed)
```

### Create dummy variable

```{r}

# create 虛擬變量
z1 <- ifelse (bfeed$race == 2, 1, 0)
z2 <- ifelse (bfeed$race == 3, 1, 0)

# new data
bfeed$z1 = z1
bfeed$z2 = z2
head(bfeed)

```


## (1) 
Write down the regression model <br>
(what is the response variable and what are the covariates):<br>

+ **$Y_i = ln(X_i) = \alpha + \gamma_1 z_{i1} + \gamma_2 z_{i2} + \gamma_3 z_{i3} + \gamma_4 z_{i4} + \sigma W_i$**
+ response variable: Duration、delta
+ covariates: race、poverty、yschool


+ Model:
+ $ln(Duration,delta) = \alpha + \gamma_1 z_{i1} + \gamma_2 z_{i2} + \gamma_3 \times poverty + \gamma_4 \times yschool + \sigma W_i$

|     |white|black|other|
|:---:|:---:|:---:|:---:|
|$z_{i1}$| 0 | 1  |  0  |
|$z_{i2}$| 0 | 0  |  1  |

```{r}
fit = survreg(Surv(duration,delta) ~ z1 + z2 + poverty + yschool, data = bfeed, dist='weibull')
summary(fit)
objects(fit)
```

+ Model:
+ $ln(Duration) = 2.2068 -0.1442 \times z_{i1} -0.2605  \times z_{i2} + 0.1914 \times poverty + 0.0510 \times yschool$

## (2) 
Estimate the regression coefficients <br>
and its corresponding 95% confidence intervals: <br><br>

Explain the meanings of your coefficient estimates in terms of <br>
how they change the baseline survival functions.

```{r warning=FALSE}
z = qnorm(0.975,lower.tail = TRUE)  # 1.959964
sqrt(fit$var)
```

### Estimate & 95% confidence intervals

```{r}

# z1: -0.1442

c(-0.1442 + z * 0.10521708, -0.1442 - z * 0.10521708)

# z2: -0.2605

c(-0.2605 + z * 0.09624615, -0.2605 - z * 0.09624615)

# poverty: 0.1914
c(0.1914 + z * 0.09391336, 0.1914 - z * 0.09391336)

# yschool: 0.0510
c(0.0510 + z * 0.019158344, 0.0510 - z * 0.019158344)

```

### Explain how they change the baseline survival functions

+ z1: -0.1442
+ 95% confidence intervals: (0.06202169, -0.35042169)

當 z1 從 0 變成 1 時(黑人媽媽)，餵母乳持續時間對數的平均將下降 0.1442 單位<br>

+ z2: -0.2605
+ 95% confidence intervals: (-0.07186101, -0.44913899)

當 z2 從 0 變成 1 時(其他種族媽媽)，餵母乳持續時間對數的平均將下降 0.2605 單位<br>

+ poverty: 0.1914
+ 95% confidence intervals: (0.375466803, 0.007333197)

當 poverty 從 0 變成 1 時(貧窮的媽媽)，餵母乳持續時間對數的平均將上升 0.1914 單位<br>

+ yschool: 0.0510
+ 95% confidence intervals: (0.08854966, 0.01345034)

當 yschool 上升 1 單位時(每多讀一年書)，餵母乳持續時間對數的平均將上升 0.0510 單位

---

+ **設 significant level = 0.05**
```{r}
z = qnorm(0.975,lower.tail = TRUE)  # 1.959964
```

## (3) 
Test whether poverty has significant effect on duration of breast feeding.

```{r}

# 藉由信賴區間:(0.375466803, 0.007333197) 可知貧窮對餵母乳時間有顯著影響
0.1914/0.09391336                        # Z = 2.038049 大於 1.959964，reject H0
pnorm(0.1914/0.09391336, lower.tail = F) # p-value : 0.02077253 小於 0.05，reject H0

```

## (4) 
Test whether mother’s education level has significant effect on during of breast feeding.

```{r}

# 藉由信賴區間:(0.08854966, 0.01345034) 可知教育年限對餵母乳時間有顯著影響
0.0510/0.019158344                        # Z = 2.662025 大於 1.959964，reject H0
pnorm(0.0510/0.019158344, lower.tail = F) # p-value : 0.0038836 小於 0.05，reject H0

```

## (5)
Test whether race has significant effect on duration of breast feeding <br>
(Use the likelihood ratio test)

### reduce model : r1 = r2 = 0

```{r}
fit2 = survreg(Surv(duration,delta) ~ poverty + yschool, data = bfeed, dist='weibull')
fit2$loglik  # Loglik(model)= -3403.5 on 2 degrees of freedom
```

### full model

```{r }
fit$loglik   # Loglik(model)= -3399.6 on 4 degrees of freedom
```

### likelihood ratio test 

```{r}
# 新:-3403.5    H0
# 原:-3399.626  H1

h0 = -3403.5
h1 = -3399.626

# 漸進 卡方 4-2
-2*(h0-h1)
qchisq(0.95, 4-2)

#  7.748 > 5.991465，拒絕H0，有 95% 顯著水準，種族跟餵母乳時間有關係

```

---
