---
title: "存活分析 HW3"
author: "B082040005 高念慈"
date: "`r Sys.Date()`"
output: html_document
---

## HW from 125~

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=F, results='hide', message=FALSE, warning=FALSE}
# install.packages('survival')
# install.packages('KMsurv')
# install.packages("fastDummies")
# library(fastDummies)           # dummy variable

library(survival)                # survival function
library(KMsurv)                  # 資料集
# install.packages("icenReg") 
# library(icenReg)               # interval censore regression
```

```{r}
data(larynx)
```

### cox model

```{r}
Fit = coxph(formula = Surv(time, delta) ~ factor(stage), data = larynx)
summary(Fit)
```

+ K-M estimate

```{r}
Fit2 = survfit(formula = Surv(time, delta) ~ factor(stage), data = larynx)
summary(Fit2)
plot(Fit2)
```

### Cox baseline cumulative hazard function

```{r}  
cum_hazard <- basehaz(Fit)
plot(cum_hazard$time, cum_hazard$hazard)
```

## Global

```{r}
Fit2 = coxph(formula = Surv(time, delta) ~ factor(stage) + age, data = larynx)
summary(Fit2)
```

## Local

### Wald test:

```{r}
Fit2 = coxph(formula = Surv(time, delta) ~ factor(stage) + age, data = larynx)

beta1hat = Fit2$coefficients[1:3]
beta10 = rep(0,3)

var11 = Fit2$var[1:3,1:3]

chiWald = t(beta1hat-beta10)%*%solve(var11)%*%(beta1hat-beta10)
chiWald

1-pchisq(chiWald,3) # pvalue:0.0004566683，拒絕H0，係數不顯著為0
```

### Likelihood ratio test:

+ reduced model(stage係數為0)

```{r}
Fit.reduced = coxph(Surv(time,delta) ~ age, data=larynx) 
summary(Fit.reduced)
```

```{r}
Fit2$loglik           # Full
Fit.reduced$loglik    # Reduce

logLik(Fit2)          # df=4
logLik(Fit.reduced)   # df=1

chiLR = 2*(Fit$loglik[2] - Fit.reduced$loglik[2])
1-pchisq(chiLR,3)     # p-value:0.003110812，拒絕H0，係數不顯著為0 # df:4-1
```

### Score test: 

+ under H0 時的係數估計

```{r}
Fit0 = coxph(formula = Surv(time, delta) ~ factor(stage) + age,
             data = larynx,
             init = c(0,0,0,Fit.reduced$coefficients), iter=0)
summary(Fit0)
```

+ score.vector

```{r}
objects(Fit0)
Fit0$score
```

```{r}
score.vector = colSums(coxph.detail(Fit0)$score)
score.vector       # -2.457351   3.089920   7.480648   5.386351e-10 

# Fit0$var

chiSC = t(score.vector[1:3])%*%Fit0$var[1:3,1:3]%*%score.vector[1:3]
1-pchisq(chiSC,3)  # p-value:0.0001063739，拒絕H0，係數不顯著為0
```

## HW:
The bfeed data frame has 927 rows and 10 columns.<br>
Format(格式)<br>
This data frame contains the following columns:

+ Duration : Duration of breast feeding, weeks
+ delta : Indicator of completed breast feeding (1=yes, 0=no)
+ race : Race of mother (1=white, 2=black, 3=other)
+ poverty(貧窮): Mother in poverty (1=yes, 0=no)
+ yschool: Education level of mother (years of school)


+ Covariates : yschool, poverty, race
+ Time to event random variable : duration of breast feeding
+ Right censoring indicator : delta (0: censored)

Note:<br> 
Confidence intervals (CI) below are all for 99% CI. Alpha level is 0.05 for all testing.

```{r}
data(bfeed)
head(bfeed)
```

### (1)
We would like to investigate (調查) how the covariates <br>
yschool, poverty and race would affect the risk function.

+ cox PH model

```{r}
model = coxph(formula = Surv(duration, delta) ~ yschool + poverty + factor(race), data = bfeed)
summary(model)
```

+ Write down the model 

|     |white|black|other|
|:---:|:---:|:---:|:---:|
|$race2$| 0 | 1  |  0  |
|$race3$| 0 | 0  |  1  |

+ Model
$$h(t|z) = h_0(t)e^{-0.05103\times yschool+0.8262 \times poverty+1.1413 \times race2 +1.2569 \times race3}$$

+ estimate the coefficients (beta)

$$\hat\beta +- Z_{\frac{\alpha}{2}}\times se(\hat\beta)$$

```{r warning=FALSE}
model$coefficients

z = qnorm(0.995, lower.tail = TRUE)  
z # 2.575829
sqrt(model$var)
```

+ 99% confidence intervals

```{r}
# yschool : -0.05103355
c(-0.05103355 - z*0.01868233, -0.05103355 + z*0.01868233)

# poverty : -0.19088136 
c(-0.19088136 - z*0.0922244, -0.19088136 + z*0.0922244)

# race2 : 0.13219468
c(0.13219468 - z*0.10307376, 0.13219468 + z*0.10307376)

# race3 : 0.22862381
c(0.22862381 - z*0.09435684, 0.22862381 + z*0.09435684)
```

```{r}
rbind(lower_CI = model$coef - z*sqrt(diag(model$var)),
      upper_CI = model$coef + z*sqrt(diag(model$var)))
```

### Please explain your results.

+ 有 99% 的信心水準各係數會落在各信賴區間內


+ exp(coef)

1. yschool : 0.95025

全母乳喂養的機率，每多讀一年書多0.95025倍(多兩年多$0.95025^2$倍)

2. poverty : 0.82623 

全母乳喂養的機率，貧窮比不貧窮多0.82623倍

3. race
+ race2 : 1.14133

全母乳喂養的機率，黑人比白人多1.14133倍

+ race3 : 1.25687

全母乳喂養的機率，其他人種比白人多1.25687倍

### (2)
If the mother’s years of school increases one year, how would the risk is changed. 

+ yschool:-0.05103355，exp(-0.05103355)=0.9502468

全母乳喂養的機率，每多讀一年書多 0.95025 倍(多兩年多$0.95025^2$倍)

### (3)
Calculate the RR (relative risk) for mothers in poverty relative to mothers not in poverty. Calculate the CI for the RR.

+ poverty:-0.19088136，exp(-0.19088136)=0.82623

全母乳喂養的機率，貧窮比不貧窮多0.82623倍

+ relative risk CI

$$P(e^{\hat\beta - Z_{\frac{\alpha}{2}}\times se(\hat\beta)}<e^{\beta}<e^{\hat\beta + Z_{\frac{\alpha}{2}}\times se(\hat\beta)})=0.99$$


```{r}
# poverty : -0.19088136
relative_risk = exp(-0.19088136)         
relative_risk                     # 0.8262306
```

```{r}
lower_poverty = (model$coef - z*sqrt(diag(model$var)))[2]
upper_poverty = (model$coef + z*sqrt(diag(model$var)))[2]
c(exp(lower_poverty), exp(upper_poverty))
```

### (4)
Calculate the RR for mothers whose race are black relative to mothers **whose race is white**. Calculate the CI for the RR 

```{r}
# race2 : 0.13219468
relative_risk_WB = exp(0.13219468)         
relative_risk_WB                     # 1.14133
```

+ relative risk CI

```{r}
lower_race_B = (model$coef - z*sqrt(diag(model$var)))[3]
upper_race_B = (model$coef + z*sqrt(diag(model$var)))[3]
c(exp(lower_race_B), exp(upper_race_B))
```

### (5)
Estimate the Breslow’s baseline hazard function and plot the estimate of cumulative baseline hazard function.

+ [使用 Cox 模型進行預測建模 - 所有關於基線危害](https://missingdatasolutions.rbind.io/2022/12/cox-baseline-hazard/)

### cumulative baseline hazard function

```{r}  
cum_hazard <- basehaz(model,centered = T) # model$means
plot(cum_hazard$time, cum_hazard$hazard)
```

### cumulative baseline hazard function

```{r}
fit_surv <- survfit(model)            # fit_surv$cumhaz
plot(fit_surv$time, fit_surv$cumhaz)
plot(fit_surv)

cum_hazard$hazard == fit_surv$cumhaz
```

### baseline hazard function

```{r}
hazard = data.frame(hazard = c(cum_hazard$hazard[1],diff(cum_hazard$hazard)),time = cum_hazard$time)
hazard
```

### (6)
Test whether the race is a significantly important for risk. <br>
Use Wald test, likelihood ratio test, and the score test. <br>
You have to write 

+ the null hypothesis/alternative hypothesis
+ the form of the test statistic and its distribution with the corresponding degrees of freedom under Ho
+ the rejection region
+ your conclusions.

## Local

### Wald test:

```{r}
### Global
model = coxph(formula = Surv(duration, delta) ~ yschool + poverty + factor(race), data = bfeed)
summary(model)
```

```{r}
### Local
beta1hat = model$coefficients[3:4]
beta10 = rep(0,2)

var11 = model$var[3:4,3:4]

chiWald = t(beta1hat-beta10)%*%solve(var11)%*%(beta1hat-beta10)
chiWald

1-pchisq(chiWald,2)
```

+ the null hypothesis/alternative hypothesis
$H_0:\beta^1=\beta_{10}=(0,0)\;vs\;H_a:\beta^1 \neq \beta_{10}$

+ the form of the test statistic and its distribution with the corresponding degrees of freedom under Ho
$under\; H_0\;,\chi^2_w=(\hat{\beta^1}-\beta_{10})^T(Var(\hat{\beta^1}))^{-1}(\hat{\beta^1}-\beta_{10})$~$\chi^2_2$

+ the rejection region
$\chi^2_w >\chi^2_{2,0.05}\;or\;p-value<0.05$

+ your conclusions.
+ pvalue:0.0360204 < 0.05，拒絕H0，係數不顯著為0，種族對全母乳喂養有顯著影響

### Likelihood ratio test:

+ reduced model(race係數為0)

```{r}
model.reduced = coxph(formula = Surv(duration, delta) ~ yschool + poverty, data = bfeed)
summary(model.reduced)
```

```{r}
model$loglik            # Full
model.reduced$loglik    # Reduce

logLik(model)           # df=4
logLik(model.reduced)   # df=2

chiLR = 2*(model$loglik[2] - model.reduced$loglik[2])
1-pchisq(chiLR,4-2)     # df:4-2
```

+ the null hypothesis/alternative hypothesis
$H_0:\beta^1=\beta_{10}=(0,0)\;vs\;H_a:\beta^1 \neq \beta_{10}$

+ the form of the test statistic and its distribution with the corresponding degrees of freedom under Ho
$under\; H_0\;,\chi^2_{LR}=2(LL(Full)-LL(Reduced))$~$\chi^2_{4-2}$

+ the rejection region
$\chi^2_{LR} >\chi^2_{2,0.05}\;or\;p-value<0.05$

+ your conclusions.
+ p-value:0.04005433 < 0.05，拒絕H0，係數不顯著為0，種族對全母乳喂養有顯著影響

### Score test: 

+ under H0 時的係數估計

```{r}
model0 = coxph(formula = Surv(duration, delta) ~ yschool + poverty + factor(race),
               data = bfeed,
               init = c(model.reduced$coefficients,0,0), iter=0)
summary(model0)
```

+ score.vector

```{r}
objects(model0)
model0$score
```

```{r}
score.vector = colSums(coxph.detail(model0)$score)
score.vector   # 6.004529e-10 -4.278418e-10  8.718373e+00  2.263892e+01 

# model0$var

chiSC = t(score.vector[3:4])%*%model0$var[3:4,3:4]%*%score.vector[3:4]
1-pchisq(chiSC,2) 
```


+ the null hypothesis/alternative hypothesis
$H_0:\beta^1=\beta_{10}=(0,0)\;vs\;H_a:\beta^1 \neq \beta_{10}$

+ the form of the test statistic and its distribution with the corresponding degrees of freedom under Ho
$under\; H_0\;,\chi^2_{sc} = (\frac{\partial LL}{\partial\beta^1}(\beta_{10},\hat{\beta^2}(\beta_{10})))^T(Var((\frac{\partial LL}{\partial\beta^1})(\beta_{10},\hat{\beta^2}(\beta_{10}))))^{-1}(\frac{\partial LL}{\partial\beta^1}(\beta_{10},\hat{\beta^2}(\beta_{10})))$~$\chi^2_2$

+ the rejection region
$\chi^2_{sc} >\chi^2_{2,0.05}\;or\;p-value<0.05$

+ your conclusions.
+ p-value:0.03562763 < 0.05，拒絕H0，係數不顯著為0，種族對全母乳喂養有顯著影響

---
---

## Time Dependent Covariates

+ Survival of patients on the waiting list for the Stanford heart transplant program

```{r}
data(heart)
```

```{r}
Fit_heart = coxph(Surv(start, stop, event) ~ age + transplant +  surgery, data=heart)
summary(Fit_heart)
```

```{r}
Fit_heart_zph <- cox.zph(Fit_heart) # scaled Schoenfeld residuals
plot(Fit_heart_zph)
```

+ use the Cox model with time-dependent covariates to fix the problem.

```{r}
data(larynx)
```

```{r}
Fit_larynx = coxph(Surv(time,delta) ~ factor(stage) + age, data=larynx)
summary(Fit_larynx)
```

+ cox.zph : scaled Schoenfeld residuals

```{r}
larynx_zph <- cox.zph(Fit_larynx) # scaled Schoenfeld residuals
plot(larynx_zph)
```

```{r}
Fit_larynx2 <- coxph(Surv(time,delta) ~ factor(stage) + age + tt(age),
                     data=larynx, tt=function(x, t, ...) x * log(t))
summary(Fit_larynx2)
```

+ Recall the marrow transplant data

```{r}
data(bmt)
```

```{r}
Fit = coxph(formula = Surv(t2,d3) ~ factor(group) + z2 + z10, data=bmt)
Fit_zph <- cox.zph(Fit)  ## provide scaled Schoenfeld residuals
plot(Fit_zph)
```

```{r}
Fit1 = coxph(formula = Surv(t2,d3) ~ factor(group) + z2 + z10, data=bmt)

cox.zph(Fit1, transform = "log")  # provide scaled Schoenfeld residuals
plot_bmt = cox.zph(Fit1) 
plot_bmt

plot(cox.zph(Fit1, transform = "log"))
plot(plot_bmt)
```

```{r}
fit2 <- coxph(Surv(t2, d3) ~ factor(group) + z2 + z10 + tt(z10),
              data = bmt,
              tt = function(x, t, ...) x * log(t))
summary(fit2)
```

+ model:

```{r}
# install.packages("splines") 
library('splines')
fit3 <- coxph(Surv(t2, d3) ~ factor(group) + z2 + bs(z10), data=bmt)
summary(fit3)
```

+ cox.zph test PH assumption (Shoenfeld residual)

```{r}
data(cancer)
```

```{r}
vfit <- coxph(Surv(time, status) ~ ph.karno + pat.karno + age, cancer)
zp <- cox.zph(vfit)
zp
```

```{r}
vfit30 <- coxph(Surv(time, status) ~ bs(ph.karno) + age, cancer)

vfit31 <- coxph(Surv(time, status) ~ ph.karno + tt(ph.karno) + age,
               data = cancer,
               tt = function(x, t, ...) x * log(t+20))  ## bs(karno)
```

### veteran

+ vfit <- coxph(Surv(time, status) ~ trt + factor(prior) + karno, veteran)
+ vfit


+ zp <- **cox.zph(vfit, transform = function(time) log(time +20))**
+ zp <- cox.zph(vfit)
+ plot(zp[3])
+ abline(0, 0, col=2)

### Our Model:
+ vfit3 <- coxph(Surv(time, status) ~ trt + factor(prior) + karno + tt(karno), data-veteran, tt=function(x, t, ...) x * log(t+20))

The time dependent coe_cient is estimated to beta(t) = -0.125 + 0.021 * log(t+20)

#### instruction of veteran dataset
+ https://stat.ethz.ch/R-manual/R-devel/library/survival/html/veteran.html

## Stratification (分層)

investigate (調查) the effect of gender (sex) and x (e.g. x=1 for treatment x=0 for control). <br>
If the variable gender (sex) does not satisfy the PH assumption. 

### Create the simplest test data set

```{r}
test1 = list(time = c(4,3,1,1,2,2,3),
             status = c(1,1,1,0,0,1,0),
             x = c(0,2,1,1,1,0,0),
             sex = c(0,0,0,0,1,1,1))

Fit_new = coxph(Surv(time, status) ~ x, test1, method = "breslow") # Breslow 
Fit_new
```

### cumulative baseline hazard estimator 

```{r}
ss = survfit(Fit_new)
plot(ss) 
plot(ss, fun="cumhaz")  # Breslow's cumulative baseline hazard estimator
```

### Fit a stratified model

```{r}
coxph(Surv(time, status) ~ x + strata(sex), test1)
```

### Stratify on the use of MTX which involves (涉及) fitting distinct baseline hazard rates to the two groups

```{r}
data(bmt)
```

```{r}
Fit = survfit(Surv(t2,d3) ~ z10, data=bmt)
# Fit = coxph(Surv(t2,d3) ~ z10, data-bmt)
```

+ plot estimated cumulative hazard function 
+ plot K-M estimator

```{r}
plot(Fit, fun="cumhaz") 
plot(Fit) 
```

+ test PH assumption value < 0.05.

```{r}
Fit = coxph(formula = Surv(t2,d3) ~ factor(group) + z2 + z10, data=bmt)
cox.zph(Fit)
```

```{r}
Fitl = coxph(formula = Surv(t2,d3) ~ factor(group) + z2 + strata(z10), data=bmt)
Fit1
```

### Hypothesis

```{r}
bmtsubset0 = bmt[bmt$z10==0,]
bmtsubset1 = bmt[bmt$z10==1,] 

Fit_nomtx = coxph(formula = Surv(t2,d3) ~ factor(group) + z2, data=bmtsubset0)
Fit_mtx = coxph(formula = Surv(t2,d3) ~ factor(group) + z2, data=bmtsubset1)

Fit_nomtx$loglik  # -230.4
Fit_mtx$loglik    # -82.58
Fit1$loglik       # -315.91

Chisq = 2*(-82.58406 - 230.4 + 315.91)  # df = 6-3 = 3

# Chisq = -2*(-310.89 + 223.34 + 82.58)
1 - pchisq(Chisq,3) # 0.11904, 不拒絕, strata model ok
```

%I(karno * log(time + 20)),

## time-dependent and time-varying Cox Model
+ https://cran.r-project.org/web/packages/survival/vignettes/timedep.pdf

```{r}
Fitl = coxph(formula = Surv(t2,d3) ~ factor(group) + z2 + z10 + I(z10*log(t2)), data=bmt)
Fitl
```

```{r}
# install.packages("splines")
# library("splines")
Fit_spline = coxph(formula = Surv(t2,d3) ~ factor(group) + z2 + bs(z10), data=bmt)
Fit_spline
```

+ time dependent covariate

```{r}
# library(survival)
data(heart)
Fit_heart = coxph(formula = Surv(start, stop, event) ~ age + transplant + surgery, data=heart)
Fit_heart
```

---
---
