---
title: "存活分析_HW2(無母數估計存活函數)"
author: "B082040005 高念慈"
date: "`r Sys.Date()`"
output: html_document
---

## HW 73-150行 & 200-217行

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=F, results='hide', message=FALSE, warning=FALSE}
# install.packages('survival')
# install.packages('KMsurv')
# install.packages("fastDummies")
# library(fastDummies)           # dummy variable

library(survival)                # survival function
library(KMsurv)                  # 資料集
# install.packages("icenReg") 
library(icenReg)                 # interval censore regression
```

# K-M estimate

```{r}
data("drug6mp")
# head(drug6mp) # packages 'KMsurv'
```

```{r}
drug6mp_event = drug6mp[drug6mp$relapse==1,] 
sort(drug6mp_event$t2)           # 6  6  6  7 10 13 16 22 23 (6-MP 的事件發生時間:t1...tD)
Surv(drug6mp$t2,drug6mp$relapse) # use Surv function to create survival object，Ti=min(xi,ci)
```

```{r}
fit.km.6mp = survfit(Surv(t2,relapse)~1, data=drug6mp)
temp = summary(fit.km.6mp) 
temp
```

```{r}
plot(fit.km.6mp, conf.int=T)  # K-M estimate plot
fit.km.6mp$n.risk             # Yi...YD
```

+ [Compute a Survival Curve for Censored Data(截尾數據)](http://stat.ethz.ch/R-manual/R-patched/library/survival/html/survfit.formula.html)
+ [在 R 中為生存選擇 conf.type](https://stats.stackexchange.com/questions/361354/choosing-conf-type-for-survfit-in-r)

```{r}
fit.km.6mp = survfit(Surv(t2,relapse)~1, data=drug6mp, conf.type="log-log")
summary(fit.km.6mp)  # conf.type=c("log","log-log","plain","none") 
```

type 方法 <br>
+ “kaplan-meier”，相當於 stype=1、ctype=1
+ “fleming-harrington”，相當於 stype=2、ctype=1 和 “fh2”

```{r}
fit.km.6mp = survfit(Surv(t2,relapse)~1, data=drug6mp, type="fh2") 
summary(fit.km.6mp)
```

+ 累積風險函數
+ $S(t)=e^{-H(t)}$

```{r}
list(temp$time,-log(temp$surv)) 
plot(temp$time,-log(temp$surv))
```

## Homework: (6mp, t1(placebo), relapse)

## (1)
Calculate K-M estimator for S(t) and variance estimate and 99% confidence interval for S(t)

```{r}
fit.km.6mp = survfit(Surv(t1,relapse)~1, data=drug6mp, ctype=1, conf.int = .99, conf.type="log-log") 
summary(fit.km.6mp)
```

+ [Be careful with standard errors in `survival::survfit`](https://dominicmagirr.github.io/post/2022-01-18-be-careful-with-standard-errors-in-survival-survfit/)

```{r warning=FALSE}
round(cbind("累積危害的標準誤差" = fit.km.6mp$std.err,
      '個別標準誤差' = summary((fit.km.6mp))$std.err,
      'variance estimate' = (summary((fit.km.6mp))$std.err)^2),4)
```

+ K-M estimator
+ variance estimate
+ 99% confidence interval

```{r warning=FALSE}
round(cbind('K-M estimator'=fit.km.6mp$surv,
            'variance estimate'=(summary((fit.km.6mp))$std.err)^2,
            '99% confidence lower'=fit.km.6mp$lower,
            '99% confidence upper'=fit.km.6mp$upper),4)
```

## (2)
Plot K-M estimators vs. time

```{r}
plot(fit.km.6mp, conf.int=F, xlab='month')  # K-M estimate plot
```

## (3)
Calculate the cumulative hazard function H(t) using the K-M estimator

+ 累積風險函數
+ $S(t)=e^{-H(t)}$
+ $-log(S(t)=H(t))$

```{r}
cbind("time" = fit.km.6mp$time,
      "H(t)" = -log(fit.km.6mp$surv)) 
plot(fit.km.6mp$time, -log(fit.km.6mp$surv), type="b")
```

## (4)
Calculate the Nelson-Aalen estimator for H(t) and S(t)

```{r warning=FALSE}

H_t = rep(0,length(fit.km.6mp$time))
s_t = rep(0,length(fit.km.6mp$time))

for (i in length(fit.km.6mp$time):1){
  H_t[i] = sum((fit.km.6mp$n.event/fit.km.6mp$n.risk)[1:i])
}

s_t = exp(-H_t)

cbind('NA_H(t)' = H_t,
      "NA_S(t)" = s_t)
```

## (5) 
Plot H(t) vs. t to check whether the distribution of the time-to-event random variable is exponential

+ Yes

```{r}
plot(fit.km.6mp$time, H_t, type="b", xlab="month", ylim=c(0,2), main="Nelson-Aalen & K-M estimator(red)", ylab='H(t)')
par(new=TRUE)
plot(fit.km.6mp$time, -log(fit.km.6mp$surv), type="b", col="red", xlab='', ylab='' , ylim=c(0,2))
```

# Left-truncated and right-censored data

```{r}
data(channing)

channing_new = channing[channing[,"ageentry"]<channing[,"age"],] # 4筆進院=出院
# only keep the subjects whose entry time (Lj) are less than the corresponding study time (Tj)

channing_new_male = channing_new[channing_new[,"gender"]==1,]
# only keep the male subjects
```

+ codes for calculating the risk sets for male subjects

```{r}
sort(channing_new_male$ageentry) # 排進院時間

sort(channing_new_male[channing_new_male[,"death"]==1,"age"]) # 排死亡時間

# cbind(channing_new$ageentry, channing_new$age)

fit = survfit(Surv(ageentry,age,death)~1, data=channing_new, subset=(gender==1))
summary(fit)
```

+ Now we only consider those deaths that occur after age 68 years old, that is P(X>t|X>68 years old) 

```{r}
fit_age = survfit(Surv(ageentry,age,death)~1, data=channing_new, subset=(gender==1 & age>=12*68))
summary(fit_age)
```

+ 以下不需要 run,但是可以看到為什麼不加 age>68 的限制時會發生的問題

```{r  results='hide', message=FALSE ,warning=FALSE}
attach(channing_new_male) # 無需實際鍵入數據框的名稱，就可使數據框中的對象可訪問
aa <- cbind(ageentry, age, death) 
## aa[order(-aa[,3], aa[,2]), ]
# aa[order(aa[,2]), ]
# aa[order(aa[,1]), ]
```

+ 一次出現兩個Model

```{r}
fit_age = survfit(Surv(ageentry,age,death)~gender, data=channing_new, subset=(age>=12*68))
summary(fit_age)
```

## HW:女性

```{r}
fit_age = survfit(Surv(ageentry,age,death)~1,
                  data=channing_new,
                  subset=(gender==2 & age>=12*68),
                  conf.type="log-log")
summary(fit_age)
```

## survival curve

```{r}
plot(fit_age$time, fit_age$surv, type='l', xlab='months', ylab='S(t)', xlim=c(800,1200), ylim=c(0,1), col='red')
par(new=T)
plot(fit_age, xlab='months', ylab='S(t)', xlim=c(800,1200), ylim=c(0,1))
```

# Interval Censoring

```{r}
left = c(0,0,6,7,7,17,37, 45,46,46)
right = c(7,8,10,16,14,Inf,44,Inf,Inf,Inf)
cbind(left, right)
```

```{r}
fitll = ic_np(cbind(left, right)) # interval censore nonparametric # S(t) 的估計
summary(fitll)
plot(fitll)
```

```{r}
fitll$T_bull_Intervals # 看区間(正的机率)總和1
fitll$p_hat            # 每個區間的機率
```

+ for practice

```{r}
left = c(5,6,7,1,13,7,19) 
right = c(9,14,8,15,15,20,Inf) 
fit22 <- ic_np(cbind(left,right)) 
fit22$T_bull_Intervals
plot(fit22)
```

# 未解釋
+ Introduction for miceData r
+ [miceData：來自 Hoel 和 Walburg 1972 的肺腫瘤間隔截尾數據](https://www.rdocumentation.org/packages/icenReg/versions/2.0.13/topics/miceData)

```{r}
data(miceData)
fit <- ic_np(cbind(1, u) ~ grp, data = miceData)  # Stratifies (分層) fits by group
plot(fit)
```

+ HW: Please estimate the survival function 
+ for the time of developing lung cancer for mice in the conventional environment 
+ and also draw the plot for the estimator.
+ 請估計小鼠在常規環境中患肺癌時間的生存函數，並為估計量畫圖

```{r}
# library('icenReg') 
# data(miceData)
# head(miceData, 3)

# np_fit = ic_np(cbind(1, u) ~ grp, data = miceData)
# summary(np_fit)
# np_fit = ic_np(cbind(1, u) ~ 1, data = miceData)

# groupedFitl <- ic_np(cbind(l,u) ~ 0, data = miceData)
# plot(groupedFitl)
# summary(groupedFitl)
```

```{r }
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
# BiocManager::install("Icens")
# install.packages('interval') 

library('BiocGenerics')
library('interval') 

id = c(1,2,3,4,5,6,7,8,9,10)
right = c(7,8,10,16,14,Inf,44,Inf,Inf,Inf)
left = c(0,0,6,7,7,17,37, 45,46,46)
k = cbind(id,left,right) 
k = data.frame(k)

plot(icfit(left,right))
fit <- icfit(Surv(left, right,type="interval2")~1, data=k)
```

+ [JSTOR](https://www.jstor.org/stable/pdf/2530965.pdf?refreqid=excelsior:4cf1715a9c4f3bf3d2 2ae5c399898e33)
+ paper: A Semiparametric Model for Regression Analysis of Interval-Censored Failure Time Data

```{r}
# install.packages("KMsurv")
# install.packages("survival")
# library(KMsurv)
# library(survival)

data(bcos)

icout <- icfit(Surv(left, right, type="interval2")~treatment, data=bcos)
# icout <- icfit(Surv(left, right, type="interval2")~1, data-bcos)
plot(icout)

## can pick out just one group
plot(icout[1])
```

---
---
